Distributed Range-Only Localisation that Preserves Sensor and Navigator Privacies

### 2021-08-16 main.tex
Included all text changes from the Automatica branch to this branch in prep for TAC resubmission (which is now approved)

\section{Introduction} \label{sec:introduction}
\IEEEPARstart{L}{ocalisation} methods in distributed sensor environments have long been an active topic of research \cite{pierceIntroductionLoran1946,ligginsDistributedDataFusion2012,liContributedReviewSourcelocalization2016} and have characterised many advancements of Kalman and Bayesian estimation theory \cite{mutambaraDecentralizedEstimationControl1998}. In particular, range-based localisation methods, including signal strength measurements \cite{wangConvexCombinationSource2018,heRangeFreeLocalizationSchemes2003}, acoustic ranges \cite{beutlerNewNonlinearFiltering2004} and ultra wideband ranges \cite{geziciLocalizationUltraWidebandRadios2005}, have found large application due to the prevalence of suitable sensors. In most cases, these localisation methods require the gathering of measurements centrally, where an estimate of location can be computed. With recent developments in distributed and cloud computing, uses of wireless and public communication channels for data transfer have becoming widespread, and the additional requirements of data privacy and state secrecy have become particularly relevant \cite{brennerSecretProgramExecution2011,renSecurityChallengesPublic2012}.

Typical cryptographic secrecy involves hiding all transferred data such that external parties in the communication network learn no new information from acquired encryptions. This can be achieved with common symmetric and public-key encryption scheme such as AES \cite{gueronIntelAdvancedEncryption2010} and RSA \cite{rivestMethodObtainingDigital1978}, respectively. In some cases however, partial data leakage or encrypted data processing is required for achieving a desired goal which has led to several homomorphic and functional encryption schemes \cite{paillierPublicKeyCryptosystemsBased1999,shiPrivacyPreservingAggregationTimeSeries2011,joyeScalableSchemePrivacyPreserving2013,chotardDecentralizedMultiClientFunctional2018} finding uses in such signal processing or localisation tasks. In \cite{alanwarPrOLocResilientLocalization2017}, homomorphic encryption is used to make time-independent model-free location estimates where an estimator does not learn sensor measurements or locations. In \cite{aristovEncryptedMultisensorInformation2018}, similar secrecy is achieved with a linear Kalman filter when a hierarchical sensor network is present. In \cite{shiPrivacyPreservingAggregationTimeSeries2011,joyeScalableSchemePrivacyPreserving2013}, privacy-preserving aggregation schemes are presented as a means to compute total powergrid usage without disclosing individual contributions, while in \cite{alexandruEncryptedCooperativeControl2019,alexandruPrivateWeightedSum2020}, centralised and hidden weighted sum aggregations, pWSAc and pWSAh, are introduced as a means for computing local control inputs in a distributed environment without learning individual inputs.

Our contribution in this work presents a range-only localisation method meeting formal cryptographic requirements that ensure sensors keep their measurements and locations private while the navigator keeps its estimates private. We first define a novel notion for private linear combination aggregation and present an implementation that satisfies these requirements, before using it to derive a filter based on the Extended Kalman Filter with no hierarchical sensor layout assumptions. The linear combination aggregation scheme we put forward is in principle similar to the pWSAh scheme in \cite{alexandruPrivateWeightedSum2020}, however, a formal definition with different communication assumptions and leakages is given, crucial for its formal security. To the best of the authors' knowledge, no existing method for Bayesian state estimation using range-only sensors and meeting the desired privacy requirements exists.

@@

We are now ready to define the type of encryption scheme we want for the specified communication protocol and the security guarantees it should provide. We let a linear combination aggregation scheme be defined as a tuple of the four algorithms $(\mathsf{Setup}, \mathsf{Enc}, \mathsf{CombEnc}, \mathsf{AggDec})$. These will be used by a trusted setup party, the navigator and sensors $i\in\{1,\dots,n\}$. They are defined as follows.
\begin{LaTeXdescription}
    \item[$\mathsf{Setup}(\kappa)$] On input of security parameter $\kappa$, generate public parameters $\mathsf{pub}$, the number of weights $m$, the navigator's public and private keys $pk_0$ and $sk_0$ and the sensor private keys $sk_i,\,i\in\{1,\dots,n\}$.
    \item[$\mathsf{Enc}(pk_0, x)$] The navigator and sensors can encrypt any value $x$ with the navigator's public key $pk_0$ and obtain the encryption $\mathcal{E}_{pk_0}(x)$.
    \item[$\mathsf{CombEnc}(t, pk_0, sk_i, \mathcal{E}_{pk_0}(\omega_1^{(t)}),\shrtdots,\mathcal{E}_{pk_0}(\omega_m^{(t)}), a^{(t)}_{i,1},\shrtdots,a^{(t)}_{i,m})$] At instance $t$, sensor $i$ computes and obtains the encrypted linear combination $l^{(t)}_i = \mathcal{E}_{pk_0,sk_i}(\sum^m_{j=1}a^{(t)}_{i,j}\omega^{(t)}_j)$ using its secret key $sk_i$.
    \item[$\mathsf{AggDec}(t, pk_0, sk_0, l^{(t)}_1,\shrtdots,l^{(t)}_n)$] At instance $t$, the navigator computes the aggregation of linear combinations $\sum^{n}_{i=1}l_i^{(t)}=\sum^{n}_{i=1}\sum^{m}_{j=1} a^{(t)}_{i,j}\omega^{(t)}_j$ using its public and private keys $pk_0$, $sk_0$.
\end{LaTeXdescription}
The security notions we want these algorithms to meet reflect the previously stated estimation privacy goals. The navigator should learn no information from individual sensors while sensors should learn no information from the navigator or any other sensors. In the context of the introduced communication protocol, this can be summarised as the following notions.
\begin{LaTeXdescription}
    \item[Indistinguishable Weights] No colluding subset of sensors gains any new knowledge about the navigator weights $\omega^{(t)}_j,\,j\in\{1,\dots,m\}$ when receiving only their encryptions from the current and previous instances and having the ability to encrypt plaintexts of their choice.
    \item[Linear Combination Aggregator Obliviousness] No colluding subset \textit{excluding} the navigator gains additional information about the remaining sensor values to be weighted, $a^{(t)}_{i,j},\,j\in\{1,\dots,m\}$, where sensor $i$ is not colluding, given only encryptions of their linear combinations $l_i$ from the current and previous instances. Any colluding subset \textit{including} the navigator learns only the sum of all linear combinations weighted by weights of their choice, $\sum^{n}_{i=1}l_i^{(t)}=\sum^{n}_{i=1}\sum^{m}_{j=1} a^{(t)}_{i,j}\omega^{(t)}_j$.
\end{LaTeXdescription}
While indistinguishable weights can be achieved by encrypting weights with an encryption scheme meeting the notion of Indistinguishability under the Chosen Plaintext Attack (IND-CPA) \cite{katzIntroductionModernCryptography2008}, the novel notion of Linear Combination Aggregator Obliviousness (LCAO) has been formalised as a typical cryptographic game between attacker and challenger in appendix \ref{app:lcao}. Lastly, we conclude the cryptographic problem definition with the following important remark.
\begin{remark}
    A leakage function including weights from the navigator requires extra care to be taken when giving its definition. If an attacker compromises the navigator, they have control over the weights, and therefore the leakage function. We note that in the leakage function above, $\sum^n_{i=1}\sum^m_{j=1}a^{(t)}_{i,j}\omega^{(t)}_j$, an individual sum weighted by the same weight may be learned by an attacker, \textit{e.g.}, $\sum^n_{i=1}a^{(t)}_{i,1}$ given weights $(1,0,\dots,0)$, but that individual sensor values $a^{(t)}_{i,j}$ remain private due to the assumption of a consistent broadcast.
\end{remark}

@@

As aggregation is typically performed on scalar inputs, we extend our notation to the context of multidimensional estimation data by letting an instance $t_{k,\tau}$ uniquely capture the scalar aggregation during an estimation timestep $k$ for a single element with position index $\tau$. To achieve this in practice, any injective function can be used, such as the concatenation $t_{k,\tau}=k\mathbin\|\tau$. The four algorithms defining our scheme are given as follows.
\begin{LaTeXdescription}
    \item[$\mathsf{Setup}(\kappa)$] On input parameter $\kappa$, generate two equal length, sufficiently large, primes $p$ and $q$, and compute $N=pq$. Define a hash function $H:\mathbb{Z} \rightarrow \mathbb{Z}_{N^2}^*$, choose a the number of weights to combine, $m>1$, and set public parameter $\mathsf{pub}=H$, navigator public key $pk_0 = N$ and navigator private key $sk_0=(p,q)$. Sensor secret keys are generated by choosing $sk_i,\,i\in\{1,\dots,n-1\}$ uniformly from $\mathbb{Z}_{N^2}$ and setting the last key as $sk_n = -\sum^{n-1}_{i=1}sk_i \pmod{N^2}$.

@@

    \item[$\mathsf{CombEnc}(t_{k,\tau}, pk_0, sk_i, \mathcal{E}_{pk_0}(\omega_j^{(k,\tau)}), a^{(k,\tau)}_{i,j}),\ j\in\{1\dots m\}$] Encrypted linear combination at instance $t_{k,\tau}$ is computed as 
    \begin{equation}
        l^{(k,\tau)}_i = H(t_{k,\tau})^{sk_i}\prod^{m}_{j=1}\mathcal{E}_{pk_0}(\omega^{(k,\tau)}_j)^{a^{(k,\tau)}_{i,j}} \pmod{N^2}\,,\label{eqn:our_scheme_lin_comb}
    \end{equation}
    and makes use of the homomorphic property \eqref{eqn:paillier_hom_mult}. Correctness follows from
    \begin{equation*}
        \begin{split}
            l^{(k,\tau)}_i &= H(t_{k,\tau})^{sk_i}\prod^{m}_{j=1}\mathcal{E}_{pk_0}(\omega^{(k,\tau)}_j)^{a^{(k,\tau)}_{i,j}} \pmod{N^2} \\
            &= H(t_{k,\tau})^{sk_i}\prod^{m}_{j=1}\mathcal{E}_{pk_0}(a^{(k,\tau)}_{i,j}\omega^{(k,\tau)}_j) \pmod{N^2} \\
            &= H(t_{k,\tau})^{sk_i}\cdot \\
            &\qquad \prod^{m}_{j=1}(N+1)^{a^{(k,\tau)}_{i,j}\omega^{(k,\tau)}_j} \rho^{N}_{j} \pmod{N^2} \\
            &= H(t_{k,\tau})^{sk_i}\cdot \\
            &\qquad (N+1)^{\sum^{m}_{j=1}a^{(k,\tau)}_{i,j}\omega^{(k,\tau)}_j} \rho_{i}^{N} \pmod{N^2}\,,
        \end{split}
    \end{equation*}
    for some values $\rho_i,\rho_j \in \mathbb{Z}_N,\,j\in\{1,\dots,m\}$. Here, $\rho_i^N$ and $H(t_{k,\tau})^{sk_i}$ can be considered the noise terms corresponding to the two levels of encryption from $pk_0$ and $sk_i$, respectively.

    \item[$\mathsf{AggDec}(t_{k,\tau}, pk_0, sk_0, l^{(k,\tau)}_1,\shrtdots,l^{(k,\tau)}_n)$] Aggregation is computed as $l^{(k,\tau)} = \prod^n_{i=1}l^{(k,\tau)}_i \pmod{N^2}$, removing the aggregation noise terms, and is followed by Paillier scheme decryption
    \begin{equation}
        \begin{split}
            \sum^{n}_{i=1}\sum^{m}_{j=1}&a^{(k,\tau)}_{i,j}\omega^{(k,\tau)}_j =\\
            &\frac{L((l^{(k,\tau)})^\lambda\pmod{N^2})}{L((N+1)^\lambda\pmod{N^2})} \pmod{N}\,,
        \end{split} \label{eqn:our_scheme_decrypt}
    \end{equation}
    with $\lambda = \mathsf{lcm}(p-1, q-1)$ and $L(u) = \frac{u-1}{N}$. The correctness of the aggregation can be seen from
    \begin{align*}
        \begin{split}
        l^{(k,\tau)} &= \prod^n_{i=1}H(t_{k,\tau})^{sk_i}\cdot \\
        &\qquad (N+1)^{\sum^{m}_{j=1}a^{(k,\tau)}_{i,j}\omega^{(k,\tau)}_j}\rho_i^N \pmod{N^2}
        \end{split}\\
        \begin{split}
            &= H(t_{k,\tau})^{\sum^n_{i=1}sk_i}\cdot \\
            &\qquad \prod^n_{i=1}(N+1)^{\sum^{m}_{j=1}a^{(k,\tau)}_{i,j}\omega^{(k,\tau)}_j}\rho_i^N \pmod{N^2}
        \end{split}\\
        &= (N+1)^{\sum^n_{i=1}\sum^{m}_{j=1}a^{(k,\tau)}_{i,j}\omega^{(k,\tau)}_j}\rho'^N \pmod{N^2}\,,
    \end{align*}
    for some values $\rho_i,\rho' \in \mathbb{Z}_N,\,i\in\{1,\dots,n\}$.
\end{LaTeXdescription}
Additionally, we note that in the above construction, all weights $\omega^{(k,\tau)}_j$ and values $a^{(k,\tau)}_{i,j}$ are integers and the resulting linear combinations and aggregation are computed modulo $N$.

@@

    \alpha_i^{(k,3)} &= (4r_{k,i}^{\prime-1})\hat{x}_{k|k-1}^2 + (-8r_{k,i}^{\prime-1}s_{x,i})\hat{x}_{k|k-1} + (4r_{k,i}^{\prime-1}s_{x,i}^2)\,, \\

@@

    \alpha_i^{(k,6)} &= (4r_{k,i}^{\prime-1})\hat{y}_{k|k-1}^2 + (-8r_{k,i}^{\prime-1}s_{y,i})\hat{y}_{k|k-1} + (4r_{k,i}^{\prime-1}s_{y,i}^2)\,.

@@

We have considered multiple sensor layouts, each with four sensors, to capture the dependence of estimated modified measurement variances $r'_{k,i}$ on the original measurements $z_{k,i}$. These layouts of varying sensor distances are shown next to the simulation initial state and a sample track in figure \ref{fig:sim_layouts}.
\begin{figure}[htbp]
    \centering
    \includegraphics{images/layouts.pdf}
    \caption{Different simulation layouts with varying distances between navigator and sensors.}
    \label{fig:sim_layouts}
\end{figure}
To demonstrate the accuracy of the method, we have compared the root mean square error (RMSE) of the privacy-preserving filter to the standard EIF using unmodified measurements, which is algebraically equivalent to the EKF typically used in industry for linearising non-linear state estimation. Estimation in each layout from figure \ref{fig:sim_layouts} consisted of $50$ filter iterations and was run $1000$ times. Unmodified measurement variances were taken as $r_{k,i}=5$ for all $k>0$ and a large fractional precision factor, $\phi=2^{32}$, was chosen. The results can be seen in figure \ref{fig:sim_layout_errors}.
\begin{figure}[htbp]
    \centering
    \includegraphics{images/layout_errors.pdf}
    \caption{Average RMSE of our privacy-preserving filter and the standard EIF for different layouts.}
    \vspace{-\baselineskip}
    \label{fig:sim_layout_errors}
\end{figure}
From these results, we can see a strong similarity in filter performance between the privacy-preserving method and that of the traditional EIF. We can also see that the varying average distances between sensors and navigator have little impact on the differences in performance. We attribute this similarity in RMSE to the conservativeness of estimated modified measurement variances $r'_{k,i}$, resulting in few additional filter divergences, and to the high fractional precision factor, keeping computations consistent with the floating-point arithmetic of the EIF.

@@

\vspace{-1\baselineskip}
\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{images/marko_ristic}}]{Marko Ristic}
received his software engineering diploma in 2018 at the University of Melbourne, Australia. In 2019, he began work as a researcher at the Intelligent Sensor-Actuator-Systems Laboratory, Karlsruhe Institute of Technology (KIT), Germany, and since 2021, he has been pursuing a Ph.D. at the Autonomous Multisensor Systems (AMS) group, Otto von Guericke University, Germany. His research interests include encrypted and privacy-preserving signal processing, focusing on state estimation, sensor fusion, and distributed localisation.
\end{IEEEbiography}

\vspace{-1\baselineskip}
\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{images/benjamin_noack}}]{Benjamin Noack}
is a professor of Computer Science at the Otto von Guericke University Magdeburg in Germany and head of the Autonomous Multisensor Systems (AMS) group. He received his diploma in computer science from the Karlsruhe Institute of Technology (KIT), Germany, in 2009. Afterward, he obtained his Ph.D. in 2013 at the Intelligent Sensor–Actuator-Systems Laboratory, Karlsruhe Institute of Technology (KIT), Germany. His research interests are in the areas of multi-sensor data fusion, distributed and decentralized Kalman filtering, combined stochastic and set-membership approaches to state estimation, and event-based systems.
\end{IEEEbiography}

\vspace{-1\baselineskip}
\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{images/uwe_d_hanebeck}}]{Uwe D. Hanebeck}
is a chaired professor of Computer Science at the Karlsruhe Institute of Technology (KIT) in Germany and director of the Intelligent Sensor–Actuator-Systems Laboratory (ISAS). He obtained his Ph.D. degree in 1997 and his habilitation degree in 2003, both in Electrical Engineering from the Technical University in Munich, Germany. His research interests are in the areas of information fusion, nonlinear state estimation, stochastic modeling, system identification, and control with a strong emphasis on theory-driven approaches based on stochastic system theory and uncertainty models. He is author and coauthor of more than 500 publications in various high-ranking journals and conferences and an IEEE Fellow.
\end{IEEEbiography}

>>>>

\section{Introduction} \label{sec:introduction}
\IEEEPARstart{L}{ocalisation} methods in distributed sensor environments have long been an active topic of research \cite{pierceIntroductionLoran1946,ligginsDistributedDataFusion2012,liContributedReviewSourcelocalization2016} and have characterised many advancements of Kalman and Bayesian estimation theory \cite{mutambaraDecentralizedEstimationControl1998}. In particular, range-based localisation methods, including signal strength measurements \cite{wangConvexCombinationSource2018,heRangeFreeLocalizationSchemes2003}, acoustic ranges \cite{beutlerNewNonlinearFiltering2004} and ultra-wideband ranges \cite{geziciLocalizationUltraWidebandRadios2005}, have found large application due to the prevalence of suitable sensors. In most cases, these localisation methods require the gathering of measurements centrally, where an estimate of location can be computed. With recent developments in distributed and cloud computing, uses of wireless and public communication channels for data transfer have become widespread, and the additional requirements of data privacy and state secrecy have become particularly relevant \cite{brennerSecretProgramExecution2011,renSecurityChallengesPublic2012}.

Typical cryptographic secrecy involves hiding all transferred data such that external parties in the communication network learn no new information from acquired encryptions. This can be achieved with common symmetric and public-key encryption scheme such as AES \cite{gueronIntelAdvancedEncryption2010} and RSA \cite{rivestMethodObtainingDigital1978}, respectively. In some cases however, partial data leakage or encrypted data processing is required for achieving a desired goal which has led to several homomorphic and functional encryption schemes \cite{paillierPublicKeyCryptosystemsBased1999,shiPrivacyPreservingAggregationTimeSeries2011,joyeScalableSchemePrivacyPreserving2013,chotardDecentralizedMultiClientFunctional2018} finding uses in signal processing or localisation tasks. In \cite{alanwarPrOLocResilientLocalization2017}, homomorphic encryption is used to make time-independent model-free location estimates where an estimator does not learn sensor measurements or locations. In \cite{aristovEncryptedMultisensorInformation2018}, similar secrecy is achieved with a linear Kalman filter when a hierarchical sensor network is present. In \cite{shiPrivacyPreservingAggregationTimeSeries2011,joyeScalableSchemePrivacyPreserving2013}, privacy-preserving aggregation schemes are presented as a means to compute total powergrid usage without disclosing individual contributions, while in \cite{alexandruEncryptedCooperativeControl2019,alexandruPrivateWeightedSum2020}, centralised and hidden weighted sum aggregations, pWSAc and pWSAh, are introduced as a means for computing local control inputs in a distributed environment without learning individual inputs.

Our contribution in this work presents a range-only localisation method meeting formal cryptographic requirements that ensure sensors keep their measurements and locations private while the navigator keeps its estimates private. We first define a novel notion for private linear combination aggregation and present an implementation that satisfies these requirements, before using it to derive a filter based on the Extended Kalman Filter with no hierarchical sensor layout assumptions. The linear combination aggregation scheme we put forward is in principle similar to the pWSAh scheme in \cite{alexandruPrivateWeightedSum2020}, however, a formal definition with different communication assumptions and leakages is given, crucial for its cryptographic security. To the best of the authors' knowledge, no existing method for Bayesian state estimation using range-only sensors and meeting the desired privacy requirements exists.

@@

We are now ready to define the type of encryption scheme we want for the specified communication protocol and the security guarantees it should provide. We let a linear combination aggregation scheme be defined as a tuple of the four algorithms $(\mathsf{Setup}, \mathsf{Enc}, \mathsf{CombEnc}, \mathsf{AggDec})$. These will be used by a trusted setup party, the navigator and sensors $i\in\{1,\dots,n\}$. They are defined as follows.
\begin{LaTeXdescription}
    \item[$\mathsf{Setup}(\kappa)$] On input of security parameter $\kappa$, generate public parameters $\mathsf{pub}$, the number of weights $m$, the navigator's public and private keys $pk_0$ and $sk_0$ and the sensor private keys $sk_i,\,i\in\{1,\dots,n\}$.
    \item[$\mathsf{Enc}(pk_0, x)$] The navigator and sensors can encrypt any value $x$ with the navigator's public key $pk_0$ and obtain the encryption $\mathcal{E}_{pk_0}(x)$.
    \item[$\mathsf{CombEnc}(t, pk_0, sk_i, \mathcal{E}(\omega_1^{(t)}),\dots,\mathcal{E}(\omega_m^{(t)}), a^{(t)}_{i,1},\dots,a^{(t)}_{i,m})$] At instance $t$, sensor $i$ computes and obtains the encrypted linear combination denoted $l^{(t)}_i = \mathcal{E}_{pk_0,sk_i}(\sum^m_{j=1}a^{(t)}_{i,j}\omega^{(t)}_j)$ using its secret key $sk_i$.
    \item[$\mathsf{AggDec}(t, pk_0, sk_0, l^{(t)}_1,\dots,l^{(t)}_n)$] At instance $t$, the navigator computes the aggregation of linear combinations $\sum^{n}_{i=1}l_i^{(t)}=\sum^{n}_{i=1}\sum^{m}_{j=1} a^{(t)}_{i,j}\omega^{(t)}_j$ using its public and private keys $pk_0$, $sk_0$.
\end{LaTeXdescription}
The security notions we want these algorithms to meet reflect the previously stated estimation privacy goals. The navigator should learn no information from individual sensors while sensors should learn no information from the navigator or any other sensors. In the context of the introduced communication protocol, this can be summarised as the following notions.
\begin{LaTeXdescription}
    \item[Indistinguishable Weights] No colluding subset of sensors gains any new knowledge about the navigator weights $\omega^{(t)}_j,\,j\in\{1,\dots,m\}$ when receiving only their encryptions from the current and previous instances and having the ability to encrypt plaintexts of their choice.
    \item[Linear Combination Aggregator Obliviousness] No colluding subset \textit{excluding} the navigator gains additional information about the remaining sensor values to be weighted, $a^{(t)}_{i,j},\,j\in\{1,\dots,m\}$, where sensor $i$ is not colluding, given only encryptions of their linear combinations $l_i$ from the current and previous instances. Any colluding subset \textit{including} the navigator learns only the sum of all linear combinations weighted by weights of their choice, $\sum^{n}_{i=1}l_i^{(t)}=\sum^{n}_{i=1}\sum^{m}_{j=1} a^{(t)}_{i,j}\omega^{(t)}_j$.
\end{LaTeXdescription}
While indistinguishable weights can be achieved by encrypting weights with an encryption scheme meeting the notion of Indistinguishability under the Chosen Plaintext Attack (IND-CPA) \cite{katzIntroductionModernCryptography2008}, the novel notion of Linear Combination Aggregator Obliviousness (LCAO) has been formalised as a typical cryptographic game between attacker and challenger in appendix \ref{app:lcao}. Lastly, we conclude the cryptographic problem definition with the following important remark.
\begin{remark}
    A leakage function including weights from the navigator requires extra care to be taken when giving its definition. If an attacker compromises the navigator, they have control over the weights, and therefore the leakage function. We note that in the leakage function above, $\sum^n_{i=1}\sum^m_{j=1}a^{(t)}_{i,j}\omega^{(t)}_j$, an individual sum weighted by the same weight may be learned by an attacker, \textit{e.g.}, $\sum^n_{i=1}a^{(t)}_{i,1}$ given weights $(1,0,\dots,0)$, but that individual sensor values $a^{(t)}_{i,j}$ remain private due to the assumption of a consistent broadcast.
\end{remark}

@@

As aggregation is typically performed on scalar inputs, we extend our notation to the context of multidimensional estimation data by letting an instance $t_{k,\tau}$ uniquely capture the scalar aggregation during an estimation timestep $k$ for a single element with position index $\tau$. To achieve this in practice, any injective function can be used, such as the concatenation $t_{k,\tau}=k\mathbin\|\tau$. The four algorithms defining our scheme are given as follows.
\begin{LaTeXdescription}
    \item[$\mathsf{Setup}(\kappa)$] On input parameter $\kappa$, generate two equal length, sufficiently large, primes $p$ and $q$, and compute $N=pq$. Define a hash function $H:\mathbb{Z} \rightarrow \mathbb{Z}_{N^2}^*$, choose the number of weights to combine, $m>1$, and set public parameter $\mathsf{pub}=H$, navigator public key $pk_0 = N$ and navigator private key $sk_0=(p,q)$. Sensor secret keys are generated by choosing $sk_i,\,i\in\{1,\dots,n-1\}$ uniformly from $\mathbb{Z}_{N^2}$ and setting the last key as $sk_n = -\sum^{n-1}_{i=1}sk_i \pmod{N^2}$.

@@

    \item[$\mathsf{CombEnc}(t_{k,\tau}, pk_0, sk_i, \mathcal{E}(\omega_1^{(k,\tau)})\dots , a^{(k,\tau)}_{i,1}\dots )$] At instance $t_{k,\tau}$, encrypted linear combination is computed as 
    \begin{equation}
        l^{(k,\tau)}_i = H(t_{k,\tau})^{sk_i}\prod^{m}_{j=1}\mathcal{E}_{pk_0}(\omega^{(k,\tau)}_j)^{a^{(k,\tau)}_{i,j}} \pmod{N^2}\,,\label{eqn:our_scheme_lin_comb}
    \end{equation}
    and makes use of the homomorphic property \eqref{eqn:paillier_hom_mult}. Correctness follows from
    \begin{equation*}
        \begin{split}
            l^{(k,\tau)}_i &= H(t_{k,\tau})^{sk_i}\prod^{m}_{j=1}\mathcal{E}_{pk_0}(\omega^{(k,\tau)}_j)^{a^{(k,\tau)}_{i,j}} \pmod{N^2} \\
            &= H(t_{k,\tau})^{sk_i}\prod^{m}_{j=1}\mathcal{E}_{pk_0}(a^{(k,\tau)}_{i,j}\omega^{(k,\tau)}_j) \pmod{N^2} \\
            &= H(t_{k,\tau})^{sk_i}\cdot \\
            &\qquad \prod^{m}_{j=1}(N+1)^{a^{(k,\tau)}_{i,j}\omega^{(k,\tau)}_j} \rho^{N}_{j} \pmod{N^2} \\
            &= H(t_{k,\tau})^{sk_i}\cdot \\
            &\qquad (N+1)^{\sum^{m}_{j=1}a^{(k,\tau)}_{i,j}\omega^{(k,\tau)}_j} \rho_{i}^{N} \pmod{N^2}\,,
        \end{split}
    \end{equation*}
    for some values $\rho_i,\rho_j \in \mathbb{Z}_N,\,j\in\{1,\dots,m\}$. Here, $\rho_i^N$ and $H(t_{k,\tau})^{sk_i}$ can be considered the noise terms corresponding to the two levels of encryption from $pk_0$ and $sk_i$, respectively.

    \item[$\mathsf{AggDec}(t_{k,\tau}, pk_0, sk_0, l^{(k,\tau)}_1,\dots,l^{(k,\tau)}_n)$] Aggregation is computed as $l^{(k,\tau)} = \prod^n_{i=1}l^{(k,\tau)}_i \pmod{N^2}$, removing the aggregation noise terms, and is followed by Paillier scheme decryption
    \begin{equation}
        \begin{split}
            \sum^{n}_{i=1}\sum^{m}_{j=1}&a^{(k,\tau)}_{i,j}\omega^{(k,\tau)}_j =\\
            &\frac{L((l^{(k,\tau)})^\lambda\pmod{N^2})}{L((N+1)^\lambda\pmod{N^2})} \pmod{N}\,,
        \end{split} \label{eqn:our_scheme_decrypt}
    \end{equation}
    with $\lambda = \mathsf{lcm}(p-1, q-1)$ and $L(u) = \frac{u-1}{N}$. The correctness of the aggregation can be seen from
    \begin{align*}
        \begin{split}
        l^{(k,\tau)} &= \prod^n_{i=1}H(t_{k,\tau})^{sk_i}\cdot \\
        &\qquad (N+1)^{\sum^{m}_{j=1}a^{(k,\tau)}_{i,j}\omega^{(k,\tau)}_j}\rho_i^N \!\!\pmod{N^2}
        \end{split}\\
        \begin{split}
            &= H(t_{k,\tau})^{\sum^n_{i=1}sk_i}\cdot \\
            &\qquad \prod^n_{i=1}(N+1)^{\sum^{m}_{j=1}a^{(k,\tau)}_{i,j}\omega^{(k,\tau)}_j}\rho_i^N \!\!\pmod{N^2}
        \end{split}\\
        &= (N+1)^{\sum^n_{i=1}\sum^{m}_{j=1}a^{(k,\tau)}_{i,j}\omega^{(k,\tau)}_j}\rho'^N \!\!\pmod{N^2}\,,
    \end{align*}
    for some values $\rho_i,\rho' \in \mathbb{Z}_N,\,i\in\{1,\dots,n\}$.
\end{LaTeXdescription}
Additionally, we note that in the above construction, all weights $\omega^{(k,\tau)}_j$ and values $a^{(k,\tau)}_{i,j}$ are integers and the resulting linear combinations and aggregation are computed modulo $N$.

@@

        \alpha_i^{(k,3)} &= (4r_{k,i}^{\prime-1})\hat{x}_{k|k-1}^2 + (-8r_{k,i}^{\prime-1}s_{x,i})\hat{x}_{k|k-1} \\
        &\qquad+ (4r_{k,i}^{\prime-1}s_{x,i}^2)\,,

@@

        \alpha_i^{(k,6)} &= (4r_{k,i}^{\prime-1})\hat{y}_{k|k-1}^2 + (-8r_{k,i}^{\prime-1}s_{y,i})\hat{y}_{k|k-1} \\
        &\qquad+ (4r_{k,i}^{\prime-1}s_{y,i}^2)\,.

@@

We have considered multiple sensor layouts, each with four sensors, to capture the dependence of estimated modified measurement variances $r'_{k,i}$ on the original measurements $z_{k,i}$. These layouts of varying sensor distances are shown next to the simulation initial state and a sample track in figure \ref{fig:sim_layouts}.
\begin{figure}[htbp]
    \centering
    \includegraphics{images/layouts.pdf}
    \caption{Different simulation layouts with varying distances between navigator and sensors.}
    \label{fig:sim_layouts}
\end{figure}
To demonstrate the accuracy of the method, we have compared the root mean square error (RMSE) of the privacy-preserving filter to the standard EIF using unmodified measurements, which is algebraically equivalent to the EKF typically used in industry for linearising non-linear state estimation. Estimation in each layout from figure \ref{fig:sim_layouts} consisted of $50$ filter iterations and was run $1000$ times. Unmodified measurement variances were taken as $r_{k,i}=5$ for all $k>0$ and a large fractional precision factor, $\phi=2^{32}$, was chosen. The results can be seen in figure \ref{fig:sim_layout_errors}.
\begin{figure}[htbp]
    \centering
    \includegraphics{images/layout_errors.pdf}
    \caption{Average RMSE of our privacy-preserving filter and the standard EIF for different layouts.}
    \vspace{-\baselineskip}
    \label{fig:sim_layout_errors}
\end{figure}
From these results, we can see a strong similarity in filter performance between the privacy-preserving method and that of the traditional EIF. We can also see that the varying average distances between sensors and the navigator have little impact on the differences in performance. We attribute this similarity in RMSE to the conservativeness of estimated modified measurement variances $r'_{k,i}$, resulting in few additional filter divergences, and to the high fractional precision factor, keeping computations consistent with the floating-point arithmetic of the EIF.

@@

%\vspace{-1\baselineskip}
\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{images/marko_ristic}}]{Marko Ristic}
received his software engineering diploma in 2018 at the University of Melbourne, Australia. In 2019, he began work as a researcher at the Intelligent Sensor-Actuator-Systems Laboratory, Karlsruhe Institute of Technology (KIT), Germany, and since 2021, he has been pursuing a Ph.D. at the Autonomous Multisensor Systems (AMS) group, Otto von Guericke University, Germany. His research interests include encrypted and privacy-preserving signal processing, focusing on state estimation, sensor fusion, and distributed localisation.
\end{IEEEbiography}

%\vspace{-1\baselineskip}
\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{images/benjamin_noack}}]{Benjamin Noack}
is a professor of Computer Science at the Otto von Guericke University Magdeburg in Germany and head of the Autonomous Multisensor Systems (AMS) group. He received his diploma in computer science from the Karlsruhe Institute of Technology (KIT), Germany, in 2009. Afterward, he obtained his Ph.D. in 2013 at the Intelligent Sensor–Actuator-Systems Laboratory, Karlsruhe Institute of Technology (KIT), Germany. His research interests are in the areas of multi-sensor data fusion, distributed and decentralized Kalman filtering, combined stochastic and set-membership approaches to state estimation, and event-based systems.
\end{IEEEbiography}

%\vspace{-1\baselineskip}
\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{images/uwe_d_hanebeck}}]{Uwe D. Hanebeck}
is a chaired professor of Computer Science at the Karlsruhe Institute of Technology (KIT) in Germany and director of the Intelligent Sensor–Actuator-Systems Laboratory (ISAS). He obtained his Ph.D. degree in 1997 and his habilitation degree in 2003, both in Electrical Engineering from the Technical University in Munich, Germany. His research interests are in the areas of information fusion, nonlinear state estimation, stochastic modeling, system identification, and control with a strong emphasis on theory-driven approaches based on stochastic system theory and uncertainty models. He is author and coauthor of more than 500 publications in various high-ranking journals and conferences and an IEEE Fellow.
\end{IEEEbiography}
